---
title: "HW_spatial_modelling"
author: "Jackson Barratt Heitmann"
date: "2/22/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(vegan)
data(BCI)
## UTM Coordinates (in metres)
BCI_xy = data.frame(x = rep(seq(625754, 626654, by=100), each=5), 
                    y = rep(seq(1011569,  1011969, by=100), len=50))

plot(BCI_xy)

#colSums because species are always the columns, and sites are the rows
sr <- colSums(BCI > 0)
hist(sr, breaks=7)
boxplot(sr)
print(sr)
quantile(sr)

sr_log <- log(sr)

sr_log
quantile(sr_log)

#tells you the index of the species that satisfy the statement
which(sr_log > 1.6 & sr_log < 1.7)
which(sr_log > 3.5 & sr_log < 3.6)

rare_sp <- BCI[ , 132]

comm_sp <- BCI[ , 62]

#rare sp plot
plot(BCI_xy, cex = rare_sp / max(rare_sp))

#comm sp plot
plot(BCI_xy, cex = comm_sp / max(comm_sp))

geo_dist <- dist(BCI_xy)

rare_dist <- dist(rare_sp)

comm_dist <- dist(comm_sp)

par(mfrow = c(1,2))
plot(geo_dist, rare_dist, main = 'rare sp' )
lines(lowess(geo_dist, rare_dist), lwd = 2, col = "red")
plot(geo_dist, comm_dist, main = ' common sp')
lines(lowess(geo_dist, comm_dist), lwd = 2, col = "blue")


cor(geo_dist, rare_dist)
cor(geo_dist, comm_dist)

mantel(geo_dist, rare_dist)
mantel(geo_dist, comm_dist)

```

1. It seems like there is spatial dependence for the common but not the rare species. We can
see from the plots that there is some spatial autocorrelation for the common species, i.e. the
further away we sample the further away the common species aggregate. However, for our rare
species, it doesn't look like we have any spatial autocorrelation, evident from the flat line,
indicating that being closer to sample where we recorded a rare species doesn't make us any
more likely to record that rare species. The results of mantel corroborate our visual
evidence, for common species (r = 0.34, p = 0.001) and for the rare species (r = -0.08, 
p = 0.91). 


```{r}

require(nlme)

sp_int <- BCI$Drypetes.standleyi

sp_one <- BCI$Hirtella.triandra



sp_ids <- c("Cordia.lasiocalyx", "Hirtella.triandra",
           "Picramnia.latifolia", "Quassia.amara",
           "Tabernaemontana.arborea", "Trattinnickia.aspera", 
           "Xylopia.macrantha", "Drypetes.standleyi")

pred_sp <- BCI[ , sp_ids]
print(pred_sp)
# need to rename columns then specify in the model
# using pred_sp$a ... 
names(pred_sp) <- c('Cord', 'Hirt', 'Picra', 'Quass',
                    'Taber', 'Tratt', 'Xylo', 'Dryp')
#Model 1 - using just one species
sp_lm <- gls(Dryp ~ Hirt, data = pred_sp)

# Plotting the Variogram for Model 1
plot(Variogram(sp_lm, form = ~ Dryp + Hirt))


sp_lm_all <- gls(Dryp ~ Hirt + Cord + Picra + 
                   Quass + Taber + Tratt + Xylo, data = pred_sp)

plot(Variogram(sp_lm_all, form = ~ Dryp + Hirt + Cord + Picra + 
                   Quass + Taber + Tratt + Xylo))

plot(Variogram(sp_lm_all, resType = 'n'))

#updating the error structures
sp_exp = update(sp_lm, corr=corRatio(form = ~ Dryp + Hirt))

# theoretically this should be plotting the residuals, we
# might have to change the shape of the line

#cor test actually tells us about the correlation coefficients







```



